Implementation and comparison of register allocation to translate \lstinline!LLVM--! to x86

### 2024-01-02 article.tex
ok

This property eliminates redefinitions entirely, seeking to simplify analysis and in turn also speed up analysis.
Although, this property also 
\begin{figure}[!ht]
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[bgcolor=LightGray]{c}
extern int add(int a, int b);

int mul(int a, int b) {
  int product = 0;
  for (int i = 0; i < a; i++)
    product = add(product, b);
  return product;
}
     \end{minted}
     \caption{Multiplication function implemented in C}\label{fig:mulc}
   \end{minipage}\hfill
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[bgcolor=LightGray]{llvm}
declare i32 @add(i32, i32)

define i32 @mul (i32 %0, i32 %1) {
 %3 = icmp sgt i32 %0, 0
 br i1 %3, label %6, label %4
4:
 %5 = phi i32 [0, %2], [%9, %6]
 ret i32 %5
6:
 %7 = phi i32 [%10, %6], [0, %2]
 %8 = phi i32 [%9, %6], [0, %2]
 %9 = call i32 @add (i32 %8, i32 %1)
 %10 = add i32 %7, 1
 %11 = icmp eq i32 %10, %0
 br i1 %11, label %4, label %6
}
     \end{minted}
     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:mulll}
   \end{minipage}
   \begin{center}
   \begin{minipage}{0.83\textwidth}
     \centering
     \begin{minted}[bgcolor=LightGray]{gas}
	.text
	.globl	add               # -- Begin function add
add:                                    # @add
	movl	%esi, %eax
	addl	%edi, %eax
	retq
     \end{minted}
     \caption{Output of \mintinline{bash}{clang -S} without debug markers}\label{fig:addx86}
   \end{minipage}
   \end{center}
\end{figure}

>>>>

This property eliminates redefinitions entirely, making it easier to reason about data flow and in turn also speed up analysis.
It's important to note that SSA form itself doesn't inherently mandate immutability, despite only allowing for single assignment.
Rather, it provides two approaches to mutable variables: the phi nodes and load/store operations.

As an extension of the earlier example, consider the implementation of multiplication in Figure \ref{fig:mulc}. Translating to LLVM IR with no optimization  can be seen in Figure \ref{fig:mulll0}, which achieves mutability by allocating a variable on the stack (lines 2-5), loading the value on use and storing on reassignment.
This means the new value is loaded and assigned upon reentering a block, so any changes made at runtime by storing new values will take effect.

%Although, this property also However, it's important to note that SSA form itself doesn't inherently introduce mutability;
\begin{figure}[!ht]
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[bgcolor=LightGray]{c}
extern int add(int a, int b);

int mul(int a, int b) {
  int product = 0;

  for (int i = 0; i < a; i++) {
    product = add(product, b);
  }

  return product;
}
     \end{minted}
     \caption{Multiplication function implemented in C}\label{fig:mulc}
     \begin{minted}[bgcolor=LightGray]{llvm}
declare i32 @add(i32, i32)

define i32 @mul (i32 %0, i32 %1) {
 %3 = icmp sgt i32 %0, 0
 br i1 %3, label %6, label %4
4:
 %5 = phi i32 [0, %2], [%9, %6]
 ret i32 %5
6:
 %7 = phi i32 [%10, %6], [0, %2]
 %8 = phi i32 [%9, %6], [0, %2]
 %9 = call i32 @add (i32 %8, i32 %1)
 %10 = add i32 %7, 1
 %11 = icmp eq i32 %10, %0
 br i1 %11, label %4, label %6
}
     \end{minted}
     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:mulll1}
   \end{minipage}\hfill
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[bgcolor=LightGray]{llvm}
define i32 @mul (i32 %0, i32 %1) {
 %3 = alloca i32
 %4 = alloca i32
 %5 = alloca i32
 %6 = alloca i32
 store i32 %0, i32* %3
 store i32 %1, i32* %4
 store i32 0, i32* %5
 store i32 0, i32* %6
 br label %7
7:
 %8 = load i32, i32* %6
 %9 = load i32, i32* %3
 %10 = icmp slt i32 %8, %9
 br i1 %10, label %11, label %18
11:
 %12 = load i32, i32* %5
 %13 = load i32, i32* %4
 %14 = call i32 @add (i32 %12, i32 %13)
 store i32 %14, i32* %5
 br label %15
15:
 %16 = load i32, i32* %6
 %17 = add i32 %16, 1
 store i32 %17, i32* %6
 br label %7
18:
 %19 = load i32, i32* %5
 ret i32 %19
}
     \end{minted}
     \caption{Stripped \texttt{clang -O0 -S -emit-llvm}}\label{fig:mulll0}
   \end{minipage}
\end{figure}
